---
title: "simpleRegress"
author: "Jasmine Walters"
date: "Spring 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r include=FALSE}
require(tidyverse)
require(tigerstats)
```
## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```
## Model WEIGHT ~ SEX

```{r}
plot(WEIGHT~SEX,data=bears)
```
```{r}
favstats(WEIGHT~SEX,data=bears)
```
The average weight of the #1 bears are about 200 lbs while the average weight of the #2 bears are about 150lbs. Group 1 on average weights 50 lbs more than group 2.



```{r}
t.test(WEIGHT~SEX,data=bears)
```

Our t-test gives us a p-value of 0,252 which is not significant since it is >0.05. This means we accept the null hypothesis. 

## Model CHEST ~ SEX

```{r}
plot(CHEST~SEX,data=bears)
```
We can see that the chest size for the #2 bears are a lot smaller than the chest size for the #1 bears. 

```{r}
favstats(CHEST~SEX,data=bears)
```

 The average chest size for #1 bears is about 37 while the average chest size for the #2 bears is about 34.
 
```{r}
t.test(CHEST~SEX,data=bears)
```
The t-test gives us a p-value of 0.1989 which is >0.05 so we are accepting the null hypothesis. There is no difference in chest size between group 1 and 2.

## Model LENGTH ~ SEX
```{r}
plot(LENGTH~SEX,data=bears)
```
We can see that the box of group 2 is smaller than group 1 but their averages are about the same. 

```{r}
favstats(LENGTH~SEX,data=bears)
```
The average length of group 1 is about 59 while the average length of group 2 is about 57. Again, these averages are very close to one another.

```{r}
t.test(LENGTH~SEX,data=bears)
```
The t-test gives us a p-value of 0.4208 which is >0.05 so we are accepting the null hypothesis. There is no difference in length between group 1 and 2.

## Model NECK ~ SEX
```{r}
plot(NECK~SEX,data=bears)
```
The bears in group 1 have bigger necks than the bears in group 2 according to this graph. 

```{r}
favstats(NECK~SEX,data=bears)
```
The average neck size in group 1 is ~22 while the average neck size in group 2 is ~18. Group 1, on average, has bigger necks than group 2.

```{r}
t.test(NECK~SEX,data=bears)
```
This t-test gives us a p-value of 0.03271 which is <0.05 so we reject the null hypothesis. Group 1's necks are bigger than group 2's necks statistically.

## Model HEADWTH ~ SEX
```{r}
plot(HEADWTH~SEX,data=bears)
```
From this graph, the headwith of group 1 is an average larger than the headwith of group 2. 

```{r}
favstats(HEADWTH~SEX,data=bears)
```
Group 1's headwith is ~7 while group 2's is about ~6. There does not seem to be a clear difference between the two. 


```{r}
t.test(HEADWTH~SEX,data=bears)
```
The p-value is 0.03559 which is <0.05, so we will reject the null hypothesis. So,. headwith is another significant difference between group 1 and 2. The #1 bears have wider heads than the #2 bears. 

## Model HEADLEN ~ SEX
```{r}
plot(HEADLEN~SEX,data=bears)
```
The head length of group 1 seems to be larger than the head lengths of group 2, even though the averages of both groups do not seem too far off from each other.

```{r}
favstats(HEADLEN~SEX,data=bears)
```
The average head length of group 1 is ~13 while the average head length of group 2 is ~12. The difference in averages does not seem too large. 

```{r}
t.test(HEADLEN~SEX,data=bears)
```
The p-value is 0.1707 which is >0.05, so we will accept the null hypothesis. There is no difference between the lengths of the heads between group 1 and 2.


